Let's talk about Thurstone \& Co.: An information-theoretical model for comparative judgments, and its statistical translation

### 2025-04-18 _tex/paper2.tex
Built site for gh-pages

Finally, using \emph{Bayesian inference methods} to transform the
model~\ref{fig-cj16_stat} into a practical statistical tool for
analyzing paired comparison data. Bayesian inference offers three key
advantages. First, it is well-suited for estimating complex and
overparameterized models, where the number of parameters may exceed the
number of observations \citep{Baker_1998, Kim_et_al_1999}. Second, it
allows researchers to incorporate prior information, which constrains
parameter estimates within plausible bounds. This feature mitigates
common estimation issues, such as non-convergence or improper solutions,
often encountered in complex models under frequentist methods
\citep{Martin_et_al_1975, Seaman_et_al_2011}. Third, Bayesian inference
is particularly effective for drawing conclusions from small samples,
where the asymptotic properties underlying frequentist methods are less
reliable \citep{Baldwin_et_al_2013, Lambert_et_al_2006, Depaoli_2014}.
The \textbf{Declarations} section of this document provides a link to
the model code, along with an alternative specification that assumes
equal discriminal dispersions. We implemented both versions of the model
using \texttt{Stan} \citep[version 2.26.1]{Stan_2020}. Although the
implementations were successful, we do not report the results here, as
they fall outside the scope of this document.

>>>>

Finally, we use \emph{Bayesian inference methods} to convert
model~\ref{fig-cj16_stat} into a practical statistical tool for
analyzing paired comparison data. Bayesian inference offers three key
advantages for this purpose. First, it handles complex and
overparameterized models--such as the one we implement--where the number
of parameters exceeds the number of observations
\citep{Baker_1998, Kim_et_al_1999}. Second, it incorporates prior
information to constrain parameter estimates within plausible bounds,
thereby mitigating estimation issues like non-convergence or improper
solutions that often affect frequentist methods
\citep{Martin_et_al_1975, Seaman_et_al_2011}. In our implementation, we
use priors to define the error distribution, which sets the scale of the
latent variables \citep{Depaoli_2014}. Third, Bayesian inference
supports robust inferences from small samples, where the asymptotic
properties underlying frequentist methods are less reliable
\citep{Baldwin_et_al_2013, Lambert_et_al_2006, Depaoli_2014}. This
feature is particularly relevant in CJ experiments, where researchers
often collect large volumes of paired comparisons but work with
relatively small samples of judges, stimuli, and individuals to test
hypotheses.

The \textbf{Declarations} section of this document provides a link to
the model code, along with an alternative specification that assumes
equal discriminal dispersions. We implemented both versions of the model
using \texttt{Stan} \citep[version 2.26.1]{Stan_2020}. Although the
implementations were successful, we do not report the results here, as
they fall outside the scope of this document.
